Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2020-04-23 paper2_JNLS/paper.tex
NM

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height \cite{muller2005rogue, ying2011linear}. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, and certain rogue features have even been recovered by exact solutions to various wave models \cite{peregrine1983water, clarkson2017rational, chen2019periodic}.
These abnormal waves can be triggered by a variety of mechanisms, including anomalous wind-forcing \cite{kharif2008influence, toffoli2017wind}, opposing currents \cite{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \cite{heller2008refraction, white1998chance}, 
and the Benjamin-Feir deep-water modulational instability \cite{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement: when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these events occur much more frequently when surface statistics deviate from Gaussian. In this way, anomalous waves can be approached from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine, holm2019stochastic}.

@@

>>>>

Rogue waves are abnormally large surface waves, defined by oceanographers as those that exceed twice the significant wave height \cite{muller2005rogue, ying2011linear}. Though such waves were once dismissed as myth, they have now been recorded in oceans across the globe and pose a recognized threat to seagoing vessels and naval structures. Rogue waves, also variously known as freak or anomalous waves, have been observed in shallow \cite{pelinovsky2000nonlinear, gramstad2013freak}, intermediate \cite{karmpadakis2019laboratory}, and deep water \cite{dematteis2018rogue, dematteis2019experimental}, \new{ and certain features of rogue waves have been recovered by exact solutions to various wave models \cite{peregrine1983water, clarkson2017rational, chen2019periodic}}.
These abnormal waves can be triggered by a variety of mechanisms, including anomalous wind-forcing \cite{kharif2008influence, toffoli2017wind}, opposing currents \cite{garrett2009rogue, onorato2011triggering}, focusing due to variable bathymetry \cite{heller2008refraction, white1998chance}, 
and the Benjamin-Feir deep-water modulational instability \cite{benjamin1967disintegration, viotti2013emergence, cousins2015unsteady, farazmand2017reduced}.
The common tie between these mechanisms is their ability to generate non-normal statistics in the surface displacement: when governed by Gaussian statistics, the likelihood of a rogue wave is extremely low, but these events occur much more frequently when surface statistics deviate from Gaussian. In this way, anomalous waves can be approached from the broader perspective of turbulent dynamical systems \cite{sapsis2013a, sapsis2013b, sapsis2013blending, chen2016filtering, majda2016introduction, macedo2017universality, MajdaQiSIAM2018, blonigan2019extreme, guth2019machine, holm2019stochastic}.

@@

	\new{We remark that Eq.~\eqref{SkewPred} bears similarity to a formula derived by Onorato \& Suret \cite{onorato2016twenty}, in the context of {\em constant-depth} KdV, that relates variations in skewness to variations in spectral bandwidth. Indeed, since $\eta_t \approx c \eta_x$ to leading order, the term $\var(\eta_x)$ in Eq.~\eqref{SkewPred} could be (approximately) interpreted as the spectral bandwidth.
We note that an analogous formula has been derived for the nonlinear Schr{\"o}dinger equation that relates kurtosis to bandwidth variations \cite{onorato2016origin}. The present study differs in that it predicts a moment-bandwidth relation for waves propagating over {\em variable} depth, where it is the depth change that creates system disequilibrium. A small transition region connects two near-equilibrium systems, i.e. one far upstream and one far downstream. The link between these two equilibrium states is made mathematically precise by the statistical matching condition \eqref{statmatch}, which is the key ingredient needed to derive Eq.~\eqref{SkewPred}. This is in contrast to the simpler situation of waves propagating over constant depth, where spectral variations and the associated moment variations arise from out-of-equilibrium initial conditions \cite{onorato2016twenty, onorato2016origin}.}
